Using automated detections to quantify the effect of postural developments on the social information in the infant view

### 2020-08-12 writeup-paper/draft_apa_july2020.tex
latest version of ms

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Automated detection of the social information}
\author{Bria L. Long\textsuperscript{1}, Alessandro Sanchez\textsuperscript{1}, Allison M. Kraus\textsuperscript{1}, Ketan Agrawal\textsuperscript{1}, \& Michael C. Frank\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\authornote{

@@

\abstract{
Social information plays a key role in infants' early linguistic and
cognitive development. Yet infants locomotor abilities also develop
continuously as they learn to sit, crawl, stand, and walk. How do these
postural developments affect infants access to the social information
relevant for early learning? To address this question, we first created
a rich annotated dataset of naturalistic play sessions between infants
and their caregivers with sets of novel and familiar toys (N=36, 8-16
months of age), recording both egocentric and third-person video. We
then developed an automated analysis method using a pose detection model
(Zhang et al., 2017) to detect the faces and hands in the infant view,
allowing us to analyze the entirety of this dataset. We found that
infants' posture and orientation to their caregiver changed dramatically
across this age range and modulated infants access to social
information; infants who were sitting or standing with their caregiver
at a close distance tended to have the most faces and hands in their
visual field. We also applied our automated method to a second
egocentric video dataset where one-year-olds explored a large play room
with their caregivers (Franchak et al., 2017), and found convergence
across both datasets, confirming the finding that motoric developments
play a significant role in shaping the social information that infants
have access to over development, and suggesting that they may play a
role in emergence of infants' cognitive and linguistic abiltiies. We
propose that the combined use of head-mounted cameras and the
application of new computer vision techniques is a promising avenue for
understanding the statistics of infants' visual and linguistic
experience as they change over development and for re-utilizing rich
video datasets collected across different labs and diverse contexts.
}

@@

We then apply this automated method to a head-mounted camera dataset
collected in our lab to directly examine the role of postural
developments on the social information in the infant view during
language learning. Specifically, we constructed a dataset where infants
and caregivers played freely using a set of novel and familiar objects;
8, 12, and 16 months-olds (N=26; N=12 in each age group) participated to
ensure a wide range of locomotor abilities. Infants' in-the-moment
posture and orientation relative to their caregiver were hand annotated
and all play sessions were transcribed. This cross-sectional design thus
allowed us to directly examine the relative contributions of age
vs.~postural developments on children's visual access to social
information, and how the avaliability of social cues changes relative to
naming events (e.g., \enquote{Yes, you like the {[}ball!{]}}). Broadly,
we predicted that there would be differential access to social
information based on children's postural developments: crawling infants
would see fewer faces/hands because they would primarily be looking at
the ground, while walking toddlers would have access to a richer visual
landscape with greater access to the social information in their
environment.

@@

We created custom annotations to describe the child's physical posture
(i.e.~standing) and the orientation of the caregiver relative to the
child (e.g.~far away). The child's posture was categorized as being
carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations; similarily, times when the child was being carried or the
caregivers were out of the frame were marked as uncodable for caregiver
orientation. On average, posture or orientation was uncodable from 1-2
minutes of data in each child (seconds excluded from analysis for
posture;. M =94.01, SD =225.18; orientation; M =94.01, SD =225.18), and
these rates did not vary substantially with the age of the child. To
assess the reliability of these annotations, a second coder coded videos
from three different children to calculate Cohen's kappa (posture, irr =
XX, orientation, irr = XX).

@@

We next analyzed the viability of OpenPose as a hand detector. Despite
the fact that hand detection is a more computationally challenging
problem (Bambach, Crandall, Smith, \& Yu, 2017), and the fact that we
used wrist keypoints as a proxy for hands, OpenPose performed moderately
well as a hand detector (F = 0.74). OpenPose achieved relatively high
precision -- generating relatively few false positives -- but showedlow
recall on the randomly sampled frames (see Table 1). As with face
detections, when we restricted our analysis to high-confidence
detections, we found much higher precision (P = 0.95), but much lower
recall (R = 0.37) and thus overall performance (F = 0.53).

@@

\subsection{Access to social information during postural
developments}\label{access-to-social-information-during-postural-developments}

@@

The interaction between infants' posture and their caregiver's
orientation had the most dramatic effect on the social information in
view. When caregivers were behind their infants, supporting their
infants' sitting or standing positions, infants saw fewer faces. When
caregivers were relatively close to their infants, infants who were
sitting or standing had more faces in view (Face detections; infant
sitting and CG close, b=0.99, SE = 0.07, Z = 14.54, P = 0; infant
standing and CG close, b=1.36, SE = 0.08, Z = 16.95, P = 0 than infants
who were lying down/crawling (i.e.~prone). When caregivers were far away
from their infants, face detections were similarity higher (Face
detections; infant sitting and CG far, b = 0.75, SE = 0.07, Z = 10.48, P
= 0; infant standing and CG far, b = 1.40, SE = 0.09, Z = 16.16, P = 0.
Overall, we found that age did not remain a significant predictor in
accounting for the faces in view when modeling infants posture, their
orientation to their caregiver, and the interaction between these two
factors (Face detections; Age (scaled), b = 0.05, SE = 0.12, Z = 0.39, P
= 0.70).

We found a similar pattern of results for wrist detections, even though
there were fewer wrist detections overall in the dataset. Infants saw
fewer wrists when caregivers were behind their infants, supporting their
infants' sitting or standing positions vs when caregivers were
relatively closer to their infants (Hand detections; infant sitting and
CG close, b=0.31, SE = 0.08, Z = 3.72, P = 0.00; infant standing and CG
close, b=0.76, SE = 0.09, Z = 8.46, P = 0 than infants who were lying
down/crawling (i.e.~prone). Wrist detections were clearly highest when
caregivers were far away from their infants (Wrist detections; infant
sitting and CG far, b=0.13, SE = 0.10, Z = 1.27, P = 0.21; infant
standing and CG far, b=1.76, SE = 0.11, Z = 15.87, P = 0. As with faces,
age did not remain a significant predictor in these models (Wrist
detections; Age (scaled), b = 0.16, SE = 0.11, Z = 1.48, P = 0.14).

We directly examined the contributions of posture and orientation by
fitting a reduced version of the full model (Nakagawa \& Schielzeth,
2013) without their fixed effects (both models were run with the maximal
random effects structure). The fixed effects in a model with only the
age of the participants accounted for relatively little variance in the
proportion of faces (marginal \(R^2\) \textless{} 0.01) or hands in view
(marginal \(R^2\) = 0.02). However, when adding infants' posture and
orientation to their caregiver to the model (and their interaction) the
marginal \(R^2\) were higher for both faces (marginal \(R^2\) = 0.28)
and wrists (marginal \(R^2\) = 0.31).\\
Overall, these results suggest that infants' visual access to social
information is largely modulated by their posture and orientation to
their caregiver, which is in turn a function of their general locomotor
development.

@@

Face detections were numerically slightly higher around these novel
naming events relative to baseline, with similar effects across age
groups (8 m.o., \(M_{fam - baseline}\) = =0.02, 12 m.o.,
\(M_{nov - baseline}\) =0.05, 16 m.o. \(M_{nov - baseline}\) =0.02)
however, they were not higher during familiar naming events vs.~baseline
(8 m.o., \(M_{fam - baseline}\) =0.00, 12 m.o., \(M_{fam - baseline}\)
=0.00, 16 m.o. \(M_{fam - baseline}\) =0.02, Conversely, wrist
detections were higher during both familiar (8 m.o.,
\(M_{fam - baseline}\) =0.07, 12 m.o., \(M_{fam - baseline}\) =0.04, 16
m.o. \(M_{fam - baseline}\) =0.08) and novel naming events relative to
baseline across all age groups (Novel; 8 m.o., (8 m.o.,
\(M_{fam - baseline}\) =0.09, 12 m.o., \(M_{fam - baseline}\) =0.08, 16
m.o. \(M_{fam - baseline}\) =0.08). These results were confirmed by a
linear mixed-effect model with scaled aged as a fixed effect and random
intercepts for each subject (Wrist detections; familiar objects
vs.~baseline, b=0.06, SE = 0.01, t = 4.33, P = 0.00'; Novel objects
vs.~baseline, b=0.08, SE = 0.02, t = 5.21, P = 0.

Overall, these exploratory results suggest that children may tend to see
more hands around naming events. This finding is consistent with the
possibility that caregivers may change how they interact with their
infant when presenting them with objects (Gogate, Bahrick, \& Watson,
2000, Gogate, Bolzani, and Betancourt (2006)). For example, caregivers
may tend to simultaneously name objects when demonstrating their
affordances or simply when pointing to them. In turn, infants may be
sensitive to these naming events and orient their attention towards
their caregiver, consistent with other accounts positing infants'
sensitivity to social cues in early word learning environments
(Yurovsky, 2018, Yurovsky and Frank (2017), Frank, Simmons, Yurovsky,
and Pusiol (2013)).

@@

\subsection{Extension to Franchak et a.,
2017}\label{extension-to-franchak-et-a.-2017}

In the present work, we found that infant's in-the-moment posture
changed with their age, as did infants' orientation relative their
caregiver. In a related study with 12-month-olds, Franchak et al. (2017)
also found that infants' in-the-moment posture changed the proportion of
time infants spent looking at faces. Here, we sought to replicate these
findings using our automated methodology (OpenPose detections), using
the footage from their head-mounted cameras (D. A. Simon, Gordon,
Steiger, \& Gilmore, 2015) for two reasons. First, we sought to validate
our novel method, which could fail to generalize to scenes from these
more complex environments, where detecting faces and hands could
arguably be a much harder task. Second, we sought to replicate the
effects of infants' in-the-moment posture on differences in visual
access to hands in an independent dataset.

@@

First, despite having different head-mounted camera setups, the view
angle of the two head-mounted cameras were only slightly different
(\(52.2^{\circ}\) horizontal by \(42.2^{\circ}\) in Franchak et al.
(2017), here \(64^{\circ}\) horizontal by \(46^{\circ}\) vertical).
However, the environment that infants were immersed in with their
caregivers (and experimenters) was much larger and more varied than the
play room used in the present dataset, containing multiple structures
and toys in different parts of the room for infants to climb, explore,
and interact with, and infants were unconstrained and allowed to freely
wander the room. In contrast, the play room used in Study 1 was
relatively small (approximately 10' x 10') and was set-up with focused
play with the pairs of novel and familiar objects. Further, multiple
people were present during the play session in Franchak et al. (2017)--
including their caregiver and two experimenters -- whereas in Study 1
the experimenters left the room during the play session.

@@

Overall, we found convergence between our two methodologies, replicating
the main results from Franchak et al. (2017), and finding that the
proportion of faces detected was greater when infants were sitting or
standing vs.~prone. We found this result regardless of whether we used
all detections (see Figure \ref{fig:franchak}, panel A; (proportion of
frames with face detections; Prone: M =0.13\%, Sitting: M =0.20\%,
Upright: M =0.22\%) or restricted our analyses to high-confidence
detections (percentage of frames with face detections; Prone: M =0.03,
Sitting: M =0.06, Upright: M =0.05). These results were confirmed in
generalized linear-mixed models, as in the first study (Sitting
vs.~Prone, b=0.34, SE = 0.02, Z = 23.37, P = 0. Upright vs.~Prone,
b=0.38, SE = 0.02, Z = 24.01, P = 24.01).

The most noticeable difference between our analysis methods is the
overall proportion of frames with faces detected by OpenPose vs.~the
proportion of frames in which infants were judged to be looking at faces
in Franchak et al. (2017). Across the entire session in Franchak et al.
(2017), infants looked at faces on 4.7\% of frames (Franchak et al.,
2017). When we used all detections from OpenPose, we found a higher
proportion of faces: M = 0.22, or 21.80\%. When we restricted our
results to only high-confidence detections, we found M = 0.06, closer to
the original values reported by Franchak et al. (2017). However, the
above analyses on the accuracy of these two methods suggest that
high-confidence detections dramatically underestimate the number of
faces in view. Thus, we suspect that these differences are due to both
(1) the fact head-mounted eye-trackers may sometimes underestimate the
proportion of faces attended due to calibration issues, and that (2)
OpenPose may sometimes detect faces that are in view for infants but
that infants may not be foveating.

We also found that infants' in-the-moment posture also modulated the
proportion of hands that were in view (i.e., wrist detections), though
these were not originally annotated in Franchak et al. (2017) (see
Figure \ref{fig:franchak}, (proportion of frames with wrist detections;
Prone: M =0.15; Sitting: M =0.21; Upright: M =0.22). These results were
confirmed in generalized linear-mixed models, as in the first study
(Sitting vs.~Prone, b=0.27, SE = 0.01, Z = 19.09, P = 0. Upright
vs.~prone, b=0.30, SE = 0.02, Z = 19.73, P = 19.73).

@@

What social cues do infants see as they learn language, and how does
this change as they grow and start to locomote themselves in their
environments? We explored this question using video data from
head-mounted cameras from two naturalistic datasets of parent-child
interactions: a cross-sectional database of naturalistic play sessions
from 8-16 month-olds with sets of novel and familiar toys, and database
of head-mounted camera videos from one-year-olds who explored a large
play area from Franchak et al. (2017). To analyze these datasets, we
developed a novel method using a pose detection model to automate the
annotation of the social information in the infant view, here
operationalized as the presence of the faces and hands of their
caregiver; these annotations were then synced with manual annotations of
infants in-the-moment posture from third-person videos.

Despite being trained on the adult perspective, the pose detector we
used (OpenPose, Cao et al. (2017)) was able to generalize relatively
well to the infant viewpoint, achieving comparable precision and
accuracy as a face detector relative to a state-of-the-art model
optimized specifically for detecting faces in natural scenes (MTCNN, K.
Zhang et al. (2016){]}. While OpenPose had relatively low recall as a
hand detector -- missing some hands that were in the infant view -- it
made comparable rates of false alarms. In both cases, we found that
overall performance was maximized when all detections were included,
regardless of their confidence, suggesting that some low-confidence face
and hand detection still index actual faces and hands that were seen by
infants. Thus, while imperfect, we suggest that OpenPose can be applied
to infant egocentric videos for the extraction of the social information
in the infant viewpoint, reducing the burden of manual annotations and
promoting the reusability of rich video datasets for further analyses.
The use of this automated methodology allowed us to easily annotate the
entirety of our dataset -- additionally analyzing the social information
around naming events -- and to re-analyze the data from Franchak et al.
(2017) to replicate our findings in a very different kind of play
session.

Broadly, our results replicate and extend previous work, first by
showing systematic changes in infants' in-the-moment posture and their
orientation relative to their caregivers (Adolph \& Franchak, 2017);
older children spent more time standing and less time sitting, and older
infants' caregivers spent less time supporting their standing or sitting
postures. Motor development changes dramatically at the same time that
children are breaking into language learning. Using these automated
detections, we found that infants' changing posture and orientation to
their caregiver jointly shaped the amount of social information that was
in their view during one-on-one play sessions with their caregivers:
Children saw the most faces/hands when they were sitting or standing and
close to their caregiver vs.~crawling or prone. These same findings were
recapitulated in a second dataset collected by Franchak et al. (2017)
with one-year-olds: sitting and upright infants saw more faces--and
hands--than infants who were prone. Motor development appears to
modulate how infants experience their visual world and the social
information in it.

While exploratory, our results also suggest that infants saw a greater
proportion of hands during naming events, hinting that children may have
been orienting towards their caregiver when they heard labels for
objects. While this effect was not present for faces, other work
(Yoshida \& Smith, 2008; Yu \& Smith, 2013, 2017), including Franchak et
al. (2017), has found that infants spend much more time looking at the
toys vs.~their caregiver's faces during these play sessions, and
highlighted the importance of hand-following as a component of joint
attention (Yu \& Smith, 2017). Further, given that there were only two
possible referents in the room at a time----and one of them was always a
familiar category (i.e., car, kitty) -- this particular play session did
not present many opportunities where children would need to use gaze
cues to disambiguate referents.

Importantly, however, all of these findings come from observational,
in-lab datasets, posing limits on their generalizability. Future work is
needed to relate these the slices of experience captured within in-lab
play sessions and infants' everyday experiences (Clerkin, Hart, Rehg,
Yu, \& Smith, 2017; Fausey et al., 2016). More broadly, though
observational findings allow us to document developmental change and
identify potential causal pathways, they cannot confirm them. Locomotive
abilities are of course only part of a cascade of changes in infants'
abilities and experiences, and these analyzes document only a fraction
of this broader, multifaceted trajectory. As children grow and change,
the activities in which they engage with their caregivers are likely to
also vary, leading to differences in the distribution of social cues
that they experience that may not be captured in these in-lab play
sessions.

Nonetheless, we think that this approach hold promise for documenting
these developmental trajectories and for generating hypotheses about
potential causal pathways. Indeed, understanding the relationship
between different domains of developmental changes in naturalistic
contexts has been a persistent challenge for developmental psychology.
However, the field of computer vision has advanced dramatically in
recent years, creating a new generation of algorithmic tools that deal
better with noisier, more complicated datasets and extract richer
information. We hope that these new tools can now be leveraged to
examine the consequences of the changing infant perspective for
linguistic, cognitive, and social development.

>>>>

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Automated detection of social information}
\author{Bria L. Long\textsuperscript{1}, Alessandro Sanchez\textsuperscript{1}, Allison M. Kraus\textsuperscript{1}, Ketan Agrawal\textsuperscript{1}, \& Michael C. Frank\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\authornote{

@@

\abstract{
Social information plays a key role in infants' early linguistic and
cognitive development. Yet infants locomotor abilities also develop
continuously as they learn to sit, crawl, stand, and walk. How do these
postural developments affect infants access to the social information
relevant for early learning? To address this question, we first created
a rich annotated dataset of naturalistic play sessions between infants
and their caregivers with sets of novel and familiar toys (N=36, 8-16
months of age), recording both egocentric and third-person video. We
then developed an automated analysis method using a pose detection model
(Zhang et al., 2017) to detect the faces and hands in the infant view,
allowing us to analyze the entirety of this dataset. We found that
infants' posture and orientation to their caregiver changed dramatically
across this age range and modulated infants access to social
information; infants who were sitting or standing with their caregiver
at a close distance tended to have the most faces and hands in their
visual field. We also applied our automated method to a second
egocentric video dataset where one-year-olds explored a large play room
with their caregivers (Franchak et al., 2017), replicating and extending
their findings. Together, these results establish the validity of a new
method for detecting social information, confirm that motoric
developments play a significant role in shaping the social information
that infants have access to. We propose that the combined use of
head-mounted cameras and the application of new computer vision
techniques is a promising avenue for re-utilizing rich video datasets
collected across different labs and diverse contexts and understanding
the statistics of infants' visual and linguistic experience as they
change over development.
}

@@

We then apply this automated method to a head-mounted camera dataset
collected in our lab to directly examine the role of postural
developments on the social information in the infant view during
language learning. Specifically, we constructed a dataset where infants
and caregivers played freely using a set of novel and familiar objects;
8, 12, and 16 months-olds (N=36; N=12 in each age group) participated to
ensure a wide range of locomotor abilities. Infants' in-the-moment
posture and orientation relative to their caregiver were hand annotated
and all play sessions were transcribed. This cross-sectional design thus
allowed us to directly examine the relative contributions of age
vs.~postural developments on children's visual access to social
information, and how the avaliability of social cues changes relative to
naming events (e.g., \enquote{Yes, you like the {[}ball!{]}}). Broadly,
we predicted that there would be differential access to social
information based on children's postural developments: crawling infants
would see fewer faces/hands because they would primarily be looking at
the ground, while walking toddlers would have access to a richer visual
landscape with greater access to the social information in their
environment.

@@

We created custom annotations to describe the child's physical posture
(i.e.~standing) and the orientation of the caregiver relative to the
child (e.g.~far away). The child's posture was categorized as being
carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations; similarily, times when the child was being carried or the
caregivers were out of the frame were marked as uncodable for caregiver
orientation. On average, posture or orientation was uncodable from 1-2
minutes of data in each child (seconds excluded from analysis for
posture, M = 94.01s, SD = 225.18s; orientation; M = 94.01s, SD =
225.18s), and these rates did not vary substantially with the age of the
child. To assess the reliability of these annotations, a second coder
coded videos from three different children to calculate Cohen's kappa
(posture, irr = XX, orientation, irr = XX).

@@

We next analyzed the viability of OpenPose as a hand detector. Despite
the fact that hand detection is a more computationally challenging
problem (Bambach, Crandall, Smith, \& Yu, 2017), and the fact that we
used wrist keypoints as a proxy for hands, OpenPose performed moderately
well as a hand detector (F = 0.74). OpenPose achieved relatively high
precision -- generating relatively few false positives -- but showed low
recall on the randomly sampled frames (see Table 1). As with face
detections, when we restricted our analysis to high-confidence
detections, we found much higher precision (P = 0.95), but much lower
recall (R = 0.37) and thus overall performance (F = 0.53).

@@

\subsection{Study 1: Access to social information during postural
developments}\label{study-1-access-to-social-information-during-postural-developments}

@@

The interaction between infants' posture and their caregiver's
orientation had the most dramatic effect on the social information in
view. When caregivers were behind their infants, supporting their
infants' sitting or standing positions, infants saw fewer faces. When
caregivers were relatively close to their infants, infants who were
sitting or standing had more faces in view (Face detections; infant
sitting and CG close, b = 0.99, SE = 0.07, Z = 14.54, P \textless{}
0.001; infant standing and CG close, b = 1.36, SE = 0.08, Z = 16.95, P
\textless{} 0.001) than infants who were lying down/crawling
(i.e.~prone). When caregivers were far away from their infants, face
detections were similarity higher (Face detections; infant sitting and
CG far, b = 0.75, SE = 0.07, Z = 10.48, P = P \textless{} 0.001; infant
standing and CG far, b = 1.40, SE = 0.09, Z = 16.16, P = P \textless{}
0.001. Overall, we found that age did not remain a significant predictor
in accounting for the faces in view when modeling infants posture, their
orientation to their caregiver, and the interaction between these two
factors (Face detections; Age (scaled), b = 0.05, SE = 0.12, Z = 0.39, P
= P = 0.697).

We found a similar pattern of results for wrist detections, even though
there were fewer wrist detections overall in the dataset. Infants saw
fewer wrists when caregivers were behind their infants, supporting their
infants' sitting or standing positions vs when caregivers were
relatively closer to their infants (Hand detections; infant sitting and
CG close, b=0.31, SE = 0.08, Z = 3.72, P \textless{} 0.001; infant
standing and CG close, b=0.76, SE = 0.09, Z = 8.46, P \textless{} 0.001)
than infants who were lying down/crawling (i.e.~prone). Wrist detections
were highest when caregivers were far away from their infants and those
infants were standing (Wrist detections;infant standing and CG far, b =
1.76, SE = 0.11, Z = 15.87, P \textless{} 0.001). As with faces, age did
not remain a significant predictor in these models (Wrist detections;
Age (scaled), b = 0.16, SE = 0.11, Z = 1.48, P = 0.140).

We directly examined the contributions of posture and orientation by
fitting a reduced version of the full model (Nakagawa \& Schielzeth,
2013) without their fixed effects (both models were run with the maximal
random effects structure). The fixed effects in a model with only the
age of the participants accounted for relatively little variance in the
proportion of faces (marginal \(R^2\) \textless{} 0.01) or hands in view
(marginal \(R^2\) = 0.02). However, when adding infants' posture and
orientation to their caregiver to the model (and their interaction) the
marginal \(R^2\) were higher for both faces (marginal \(R^2\) = 0.28)
and wrists (marginal \(R^2\) = 0.31). Overall, these results suggest
that infants' visual access to social information is largely modulated
by their posture and orientation to their caregiver, which is in turn a
function of their general locomotor development.

@@

Face detections were numerically slightly higher around these novel
naming events relative to baseline, with similar effects across age
groups (8 m.o., \(M_{fam - baseline}\) = 0.02, 12 m.o.,
\(M_{nov - baseline}\) = 0.05, 16 m.o. \(M_{nov - baseline}\) = 0.02)
however, they were not higher during familiar naming events vs.~baseline
(8 m.o., \(M_{fam - baseline}\) =0.00, 12 m.o., \(M_{fam - baseline}\) =
0.00, 16 m.o. \(M_{fam - baseline}\) = 0.02). Conversely, wrist
detections were higher during both familiar (8 m.o.,
\(M_{fam - baseline}\) =0.07, 12 m.o., \(M_{fam - baseline}\) = 0.04, 16
m.o. \(M_{fam - baseline}\) = 0.08) and novel naming events relative to
baseline across all age groups (8 m.o., \(M_{nov - baseline}\) = 0.09,
12 m.o., \(M_{nov - baseline}\) = 0.08, 16 m.o. \(M_{nov - baseline}\) =
0.08). These results were confirmed by a linear mixed-effect model with
scaled aged as a fixed effect and random intercepts for each subject
(Wrist detections; familiar objects vs.~baseline, b = 0.06, SE = 0.01, t
= 4.33, P \textless{} 0.001; Novel objects vs.~baseline, b = 0.08, SE =
0.02, t = 5.21, P \textless{} 0.001).

Overall, these exploratory results suggest that children may tend to see
more hands around naming events. This finding is consistent with the
possibility that caregivers may change how they interact with their
infant when presenting them with objects (Gogate, Bahrick, \& Watson,
2000; Gogate, Bolzani, \& Betancourt, 2006). For example, caregivers may
tend to simultaneously name objects when demonstrating their affordances
or simply when pointing to them. In turn, infants may be sensitive to
these naming events and orient their attention towards their caregiver,
consistent with other accounts positing infants' sensitivity to social
cues in early word learning environments (Frank, Simmons, Yurovsky, \&
Pusiol, 2013; Yurovsky, 2018; Yurovsky \& Frank, 2017).

@@

\subsection{Study 2: Extension of Franchak et al.,
2017}\label{study-2-extension-of-franchak-et-al.-2017}

In the present work, we found that infant's in-the-moment posture
changed with their age, as did infants' orientation relative their
caregiver. In a related study with 12-month-olds, Franchak et al. (2017)
also found that infants' in-the-moment posture changed the proportion of
time infants spent looking at faces. Here, we sought to replicate these
findings using our automated methodology (OpenPose detections), using
the footage from their head-mounted cameras (hosted on Databrary; D. A.
Simon, Gordon, Steiger, \& Gilmore, 2015) for two reasons. First, we
sought to validate our novel method, which could fail to generalize to
scenes from these more complex environments, where detecting faces and
hands could arguably be a much harder task. Second, we sought to
replicate the effects of infants' in-the-moment posture on differences
in visual access to hands in an independent dataset.

@@

First, despite having different head-mounted camera setups, the view
angle of the two head-mounted cameras used in Study 1 and Franchak et
al. (2017) were only slightly different (\(52.2^{\circ}\) horizontal by
\(42.2^{\circ}\) in Franchak et al. (2017), here \(64^{\circ}\)
horizontal by \(46^{\circ}\) vertical). However, the play environment in
Franchak et al. (2017) that infants were immersed in with their
caregivers (and experimenters) was much larger and more varied than the
play room used in the present dataset, containing multiple structures
and toys in different parts of the room for infants to climb, explore,
and interact with, and infants were unconstrained and allowed to freely
wander the room. In contrast, the play room used in Study 1 was
relatively small (approximately 10' x 10') and was set-up with focused
play with the pairs of novel and familiar objects. Further, multiple
people were present during the play session in Franchak et al. (2017)--
including their caregiver and two experimenters -- whereas in Study 1
the experimenters left the room during the play session.

@@

Overall, we found convergence between our two methodologies, replicating
the main results from Franchak et al. (2017), and finding that the
proportion of faces detected was greater when infants were sitting or
standing vs.~prone. We found this result regardless of whether we used
all detections (see Figure \ref{fig:franchak}, panel A; (proportion of
frames with face detections; Prone: M = 0.13, Sitting: M = 0.20,
Upright: M = 0.22) or restricted our analyses to high-confidence
detections (percentage of frames with face detections; Prone: M = 0.03,
Sitting: M = 0.06, Upright: M = 0.05). These results were confirmed in
generalized linear-mixed models, as in the first study (Sitting
vs.~Prone, b = 0.34, SE = 0.02, Z = 23.37, P \textless{} 0.001. Upright
vs.~Prone, b = 0.38, SE = 0.02, Z = 24.01, P \textless{} 0.001).

The most noticeable difference between our analysis methods is the
overall proportion of frames with faces detected by OpenPose vs.~the
proportion of frames in which infants were judged to be looking at faces
in Franchak et al. (2017). Across the entire session in Franchak et al.
(2017), infants looked at faces on 4.7\% of frames (Franchak et al.,
2017). When we used all detections from OpenPose, we found a higher
proportion of faces -- 21.80\%. When we restricted our results to only
high-confidence detections, we found 6.11\% of frames with faces, closer
to the original values reported by Franchak et al. (2017). However, the
above analyses on the accuracy of this method suggest that
high-confidence detections dramatically underestimate the number of
faces in view. Thus, we suspect that these differences between our
findings are due to both (1) the fact head-mounted eye-trackers may
sometimes underestimate the proportion of faces attended due to
calibration issues, and that (2) OpenPose may sometimes detect faces
that are in view for infants but that infants may not be foveating.

We also found that infants' in-the-moment posture also modulated the
proportion of hands that were in view (i.e., wrist detections), though
these were not originally annotated in Franchak et al. (2017) (see
Figure \ref{fig:franchak}, proportion of frames with wrist detections;
Prone: M = 0.15; Sitting: M = 0.21; Upright: M = 0.22). These results
were confirmed in generalized linear-mixed models, as in the first study
(Sitting vs.~Prone, b = 0.27, SE = 0.01, Z = 19.09, P \textless{} 0.001.
Upright vs.~prone, b = 0.30, SE = 0.02, Z = 19.73, P \textless{} 0.001).

@@

What social cues do infants see as they learn language, and how does
this change as they grow and start to locomote themselves? We explored
this question using video data from head-mounted cameras from two
naturalistic datasets of parent-child interactions: a cross-sectional
database of play sessions from 8-16 month-olds with sets of novel and
familiar toys, and database of head-mounted camera videos from
one-year-olds who explored a large play area from Franchak et al.
(2017). To analyze these datasets, we developed a novel method using a
pose detection model to automate the annotation of the social
information in the infant view, here operationalized as the presence of
the faces and hands of their caregiver; these annotations were then
synced with manual annotations of infants in-the-moment posture from
third-person videos.

Despite being trained on the adult perspective, the pose detector we
used (OpenPose, Cao et al. (2017)) was able to generalize relatively
well to the infant viewpoint, achieving comparable precision and
accuracy as a face detector relative to a state-of-the-art model
optimized specifically for detecting faces in natural scenes (MTCNN, K.
Zhang et al. (2016)). While OpenPose had relatively low recall as a hand
detector -- missing some hands that were in the infant view -- it made
comparable rates of false alarms. In both cases, we found that overall
performance was maximized when all detections were included, regardless
of their confidence, suggesting that some low-confidence face and hand
detection still index actual faces and hands that were seen by infants.
Thus, while imperfect, we suggest that OpenPose can be applied to infant
egocentric videos for the extraction of the social information in the
infant viewpoint, reducing the burden of manual annotations and
promoting the reusability of rich video datasets for further analyses.
The use of this automated methodology allowed us to easily annotate the
entirety of our dataset -- additionally analyzing the social information
around naming events -- and to re-analyze the data from Franchak et al.
(2017) to replicate our findings in a very different kind of play
session.

Broadly, our results replicate and extend previous work, first by
showing systematic changes in infants' in-the-moment posture and their
orientation relative to their caregivers (Adolph \& Franchak, 2017);
older children spent more time standing and less time sitting, and older
infants' caregivers spent less time supporting their standing or sitting
postures. Motor development changes dramatically at the same time that
children are breaking into language learning. Using these automated
detections, we found that infants' changing posture and orientation to
their caregiver jointly shaped the amount of social information that was
in their view during one-on-one play sessions with their caregivers.
Children saw the most faces/hands when they were sitting or standing and
close to their caregiver vs.~crawling or prone. These same findings were
recapitulated in a second dataset collected by Franchak et al. (2017)
with one-year-olds: sitting and upright infants saw more faces -- and
hands -- than infants who were prone. Motor development appears to
modulate how infants experience their visual world and the social
information in it.

While exploratory, our results also suggest that infants saw a greater
proportion of hands around naming events, hinting that children may have
been orienting towards their caregiver when they heard labels for
objects. While this effect was not present for faces, other work
(Yoshida \& Smith, 2008; Yu \& Smith, 2013, 2017), including Franchak et
al. (2017), has found that infants spend much more time looking at the
toys vs.~their caregiver's faces during these play sessions, and
highlighted the importance of hand-following as a component of joint
attention (Yu \& Smith, 2017). Further, given that there were only two
possible referents in the room at a time -- and one of them was always a
familiar category -- this particular play session did not present many
opportunities where children would need to use gaze cues to disambiguate
referents.

Importantly, however, all of these findings come from observational,
in-lab datasets, posing important limits on their generalizability.
Future work is needed to relate these slices of experience captured
within in-lab play sessions with infants' everyday experiences (Clerkin,
Hart, Rehg, Yu, \& Smith, 2017; Fausey et al., 2016). More broadly,
though observational findings allow us to document developmental change
and identify potential causal pathways, they cannot confirm them.
Locomotive abilities are of course only part of a cascade of changes in
infants' abilities and experiences, and these analyses document only a
fraction of this broader, multifaceted trajectory. As children grow and
change, the activities in which they engage with their caregivers are
likely to also vary, leading to differences in the distribution of
social cues that they experience that may not be captured in these
in-lab play sessions.

Understanding the relationship between different domains of
developmental changes in naturalistic contexts has been a persistent
challenge for developmental psychology. Nonetheless, we think that this
approach hold promise for documenting these developmental trajectories
and for generating new hypotheses.The field of computer vision has
advanced dramatically in recent years, creating a new generation of
algorithmic tools that deal better with noisier, more complicated
datasets and extract richer information. We hope that these new tools
can now be leveraged to examine the consequences of the changing infant
perspective for linguistic, cognitive, and social development.
